---
title: "Project Proposal: What Makes a Song Popular on Spotify?"
author: "CCBK"
date: "3/21/19"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(infer)
library(broom)
songs <- read_csv("/cloud/project/data/SpotifyFeatures.csv")
```

## Section 1. Introduction

The goal of our project is to determine the qualities of a song that influence popularity on Spotify. The dataset was retrieved online at https://www.kaggle.com/edalrami/19000-spotify-songs. The data is derived from the Spotify Web API, which uses internal Spotify metrics to determine the parameters of the dataset. For subjective variables such as danceability and instrumentalness, the Spotify Web API utilizes Echo Nest algorithms. The dataset curator retrieved data for 19,000 songs at random and aggregated this information into the dataset. Below is a brief description explaining the 18 variables that are within the chosen dataset. 
 
+ Categorical Variables:
Genre, artist_name, track_name, track_id, mode, time_signature (An estimated overall time signature of a track), key.

+ Continuous Numerical Variables:
duration_ms (Duration of song in miliseconds), loudness (The overall loudness of the track in decibels), tempo (The overall estimate beats per minute of the track).

+ Discrete Numerical Variables:
popularity (The popularity of the track from 0 to 100 based on recent stream counts),
acousticness (A confidence measure from 0 to 1.0 of whether the song is acoustic),
danceability (How suitable a song is for dancing from 0 to 1.0), liveness (The likelihood that the track was performed live from 0 to 1.0), valence (A measure from 0.0 to 1.0 describing the musical positiveness), instrumentalness (A confidence measure from 0 to 1.0 on whether song is instrumental), energy (A song’s energy from 0.0 to 1.0 considering various factors), speechiness (A measure of how exclusively speech-like the track is from 0 to 1.0).

## Section 2. Data analysis plan

This project will analyze the variables of a song that correlate with a higher song popularity. To find this, we will assign song popularity as our dependent response variable and the other variables such as danceability, energy, and key as our predictor variables. Using single predictors models and backwards selection using multiple predictors models, we can determine which variables are the best predictors of a song's popularity score. Higher R^2 value for single predictors model or a higher adjusted R^2 value for multiple predictors can help us determine this. 

In addition to looking at individual songs, we intend to group songs by genres or playlists. After grouping, we will explore the difference in popularity means/medians within certain genres. For example, if the genres rap and pop each have 100 songs within the data set, we will perform simulations on samples with replacement from both genres to create a distribution of the difference in means. In this case, popularity score would be the response variable and the genre would be the explanatory variable. Then we will look at the p-value (p-value < 0.05 for proving dependence) to determine if there is a significant difference. 95% confidence intervals for the difference in popularity score means or medians can also be used to make conclusions about the difference in populations. 

Visualizations can be used to determine skewness of the data set. This will help us determine whether mean/SD or median/IQR is more fitting to describe the certain variables or groupings. We will be interested in looking at the distribution of different variables. A skew will indicate that the songs tend towards a certain type of sound. The goal is to have a representative sample of the entirety of spotify, with a range of different popularities, keys, liveliness, and more.  

## Section 3. Data

```{r glimpse}
glimpse(songs)
```
