---
title: "Linear Regression"
output: github_document
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(infer)
library(broom)
library(forcats)
songs <- read_csv("/cloud/project/data/SpotifyFeatures.csv")

glimpse(songs)
```

```{r full-model}

full_model = lm(popularity ~ acousticness +
                  danceability +
                  energy +
                  instrumentalness +
                  liveness +
                  loudness +
                  speechiness + 
                  tempo + 
                  valence +
                  energy * danceability,
                data = songs)

full_model

glance(full_model)$r.squared
glance(full_model)$adj.r.squared
```


```{r selected-model}

selected_model <- step(full_model, direction = "backward")

selected_model

glance(selected_model)$r.squared
glance(selected_model)$adj.r.squared
```

score_hat = 54.92 + -11.24 * acousticness + 16.22 * danceability - 8.11 * energy - 3.61 * instrumentalness + -10.06 * liveness + 0.71 * loudness + -9.89 * speechiness + 0.018 * tempo + -8.90 * valence + -13.36 * danceability:energy

The adjusted R squared value for this model is 0.28, which means that roughly 28.2% of the variabilit 
he R squared value for this model is 0.042, which means that roughly 4.2% of the variability in a Professor’s score can be explained by their class credits. This aligns with our suspicions, because this percentage is quite low, which suggests that class credits cannot very accurately predict score.