---
title: "Linear Regression"
output: github_document
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(infer)
library(broom)
library(forcats)
songs <- read_csv("/cloud/project/data/SpotifyFeatures.csv")

glimpse(songs)
```

```{r full-model}

full_model = lm(popularity ~ acousticness +
                  danceability +
                  energy +
                  instrumentalness +
                  liveness +
                  loudness +
                  speechiness + 
                  tempo + 
                  valence +
                  energy * danceability,
                data = songs)

full_model

glance(full_model)$r.squared
glance(full_model)$adj.r.squared
```


```{r selected-model}

selected_model <- step(full_model, direction = "backward")

selected_model

glance(selected_model)$r.squared
glance(selected_model)$adj.r.squared
```

score_hat = 54.92 + -11.24 * acousticness + 16.22 * danceability - 8.11 * energy - 3.61 * instrumentalness + -10.06 * liveness + 0.71 * loudness + -9.89 * speechiness + 0.018 * tempo + -8.90 * valence + -13.36 * danceability:energy

When all variables have a value of 0, then it is expected that a song's populariy is 54.92. One example of how a variable influences the popularity is with acousticness. If a song's acoustic score increases by 1 then popularity will decreases by 11.14 points. There is also an apparent interaction between danceability and energy. When the interaction increases by one the popularity decreases by 12.96 points. 

The adjusted R squared value for this model is 0.28, which means that roughly 28.2% of the variability in a song's popularity can be explained by the different variables in the multiple regression model. This indicates that there is a weak to moderate positive overall realtionship between a song's popularity and the different variables related to a song.
