---
title: "Permute"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(infer)
library(broom)
library(forcats)
songs <- read_csv("/cloud/project/data/songs.csv")
nrep <- 1000
```

### Popularity Difference Between Pop and Rap 

Null Hypothesis: There is not a difference between the mean popularities of songs that are pop and songs that are rap. 
Alternative Hypothesis: There is a difference between the mean popularities of songs that are pop and songs that are rap.

```{r observed-difference-pop}
songs_poprap <- songs%>%
  filter(genre=="Pop" | genre == "Rap")
mean_poprap <- songs_poprap %>%
  group_by(genre) %>%  
  summarise(samp_mean = mean(popularity)) %>%
  summarise(diff(samp_mean)) %>%
  pull()
mean_poprap
```

```{r null-dist-pop}
null_dist_poprap <- songs_poprap %>%
  specify(response = popularity, explanatory = genre) %>%
  hypothesize(null = "independence") %>%
  generate(reps = nrep, type = "permute") %>%
  calculate(stat = "diff in means", order = c("Pop", "Rap"))
ggplot(data = null_dist_poprap, aes(x = stat)) +
  geom_histogram(binwidth=.2) +
  geom_vline(xintercept = mean_poprap, color = "red") +
  geom_vline(xintercept = -1 * mean_poprap, color = "red") +
  labs(title = "Null distribution of differences in popularity means for pop or rap songs")
```

```{r p-value-pop}
null_dist_poprap %>%
  filter(stat <= mean_poprap) %>%
  summarise(pvalue = 2 * (n() / nrep))
```

The p-value of the difference between pop and rap songs is 0, which mean we can reject the null hypothesis that there is not a difference between the mean popularities of songs that are pop versus rap. The data provides convincing evidence that there is a difference in mean popularities of songs that in the pop genre versus the rap genre. 

### Popularity Difference Between High Energy and Low Energy
Null Hypothesis: There is not a difference between the mean popularities of songs that have high energy and low energy.
Alternative Hypothesis: There is a difference between the mean popularities of songs that have high energy and low energy.

```{r observed-difference-energy}
songs <- songs%>%
  mutate(energyLevel = case_when(
    energy >= .5 ~ "High",
    energy < .5 ~ "Low"
  ))
mean_energy <- songs %>%
  group_by(energyLevel) %>%  
  summarise(samp_mean = mean(popularity)) %>%
  summarise(diff(samp_mean)) %>%
  pull()
mean_energy
```

```{r null-dist-energy}
null_dist_energy <- songs %>%
  specify(response = popularity, explanatory = energyLevel) %>%
  hypothesize(null = "independence") %>%
  generate(reps = nrep, type = "permute") %>%
  calculate(stat = "diff in means", order = c("High", "Low"))
ggplot(data = null_dist_energy, aes(x = stat)) +
  geom_histogram(binwidth=.2) +
  geom_vline(xintercept = mean_energy, color = "red") +
  geom_vline(xintercept = -1 * mean_energy, color = "red") +
  labs(title = "Null distribution of differences in popularity means based on energy")
```

```{r p-value-energy}
null_dist_energy %>%
  filter(stat <= mean_energy) %>%
  summarise(pvalue = 2 * (n() / nrep))
```
The p-value of the difference between high and low energy is 0, which mean we can reject the null hypothesis that there is not a difference between the mean popularities of songs that have high energy and low energy. The data provides convincing evidence that there is a difference in mean popularities of songs that have high energy versus low energy. 

### Popularity Difference Between High Danceability and Low Danceabililty
Null Hypothesis: There is not a difference between the mean popularities of songs that have high danceability and low danceability
Alternative Hypothesis: There is a difference between the mean popularities of songs that have high danceability and low danceability

```{r observed-difference-dance}
songs <- songs%>%
  mutate(danceLevel = case_when(
    danceability >= .5 ~ "High",
    danceability < .5 ~ "Low"
  ))
mean_dance <- songs %>%
  group_by(danceLevel) %>%  
  summarise(samp_mean = mean(popularity)) %>%
  summarise(diff(samp_mean)) %>%
  pull()
mean_dance
```

```{r null-dist-dance}
null_dist_dance <- songs %>%
  specify(response = popularity, explanatory = danceLevel) %>%
  hypothesize(null = "independence") %>%
  generate(reps = nrep, type = "permute") %>%
  calculate(stat = "diff in means", order = c("High", "Low"))
ggplot(data = null_dist_dance, aes(x = stat)) +
  geom_histogram(binwidth=.2) +
  geom_vline(xintercept = mean_dance, color = "red") +
  geom_vline(xintercept = -1 * mean_dance, color = "red") +
  labs(title = "Null distribution of differences in popularity means based on danceability")
```

```{r p-value-dance}
null_dist_dance %>%
  filter(stat <= mean_dance) %>%
  summarise(pvalue = 2 * (n() / nrep))
```

The p-value of the difference between high and low danceability is 0, which means we can reject the null hypothesis that there is not a difference between the mean popularities of songs that have high danceability and low danceability The data provides convincing evidence that there is a difference in mean popularities of songs that have high danceability versus low danceability 

### Popularity Difference Between High Speechiness and Low Speechiness
Null Hypothesis: There is not a difference between the mean popularities of songs that have high speechiness and low speechiness
Alternative Hypothesis: There is a difference between the mean popularities of songs that have high speechiness and low speechiness

```{r observed-difference-speech}
songs <- songs%>%
  mutate(speechLevel = case_when(
    speechiness >= .5 ~ "High",
    speechiness < .5 ~ "Low"
  ))
mean_speech <- songs %>%
  group_by(speechLevel) %>%  
  summarise(samp_mean = mean(popularity)) %>%
  summarise(diff(samp_mean)) %>%
  pull()
mean_speech
```

```{r null-dist-speech}
null_dist_speech <- songs %>%
  specify(response = popularity, explanatory = speechLevel) %>%
  hypothesize(null = "independence") %>%
  generate(reps = nrep, type = "permute") %>%
  calculate(stat = "diff in means", order = c("High", "Low"))
ggplot(data = null_dist_speech, aes(x = stat)) +
  geom_histogram(binwidth=.2) +
  geom_vline(xintercept = mean_speech, color = "red") +
  geom_vline(xintercept = -1 * mean_speech, color = "red") +
  labs(title = "Null distribution of differences in popularity means based on speechiness")
```

The p-value of the difference between high and low speechiness is 0, which means we can reject the null hypothesis that there is not a difference between the mean popularities of songs that have high speechiness and low speechiness The data provides convincing evidence that there is a difference in mean popularities of songs that have high speechiness versus low speechiness. 

### Popularity Difference Between High Valence and Low Valence
Null Hypothesis: There is not a difference between the mean popularities of songs that have high valence and low valence
Alternative Hypothesis: There is a difference between the mean popularities of songs that have high valence and low valence

```{r observed-difference-valence}
songs <- songs%>%
  mutate(valenceLevel = case_when(
    valence >= .5 ~ "High",
    valence < .5 ~ "Low"
  ))
mean_valence <- songs %>%
  group_by(valenceLevel) %>%  
  summarise(samp_mean = mean(popularity)) %>%
  summarise(diff(samp_mean)) %>%
  pull()
mean_valence
```

```{r null-dist-valence}
null_dist_valence <- songs %>%
  specify(response = popularity, explanatory = valenceLevel) %>%
  hypothesize(null = "independence") %>%
  generate(reps = nrep, type = "permute") %>%
  calculate(stat = "diff in means", order = c("High", "Low"))
ggplot(data = null_dist_valence, aes(x = stat)) +
  geom_histogram(binwidth=.2) +
  geom_vline(xintercept = mean_valence, color = "red") +
  geom_vline(xintercept = -1 * mean_valence, color = "red") +
  labs(title = "Null distribution of differences in popularity means based on speechiness")
```

```{r p-value-valence}
null_dist_valence %>%
  filter(stat <= mean_valence) %>%
  summarise(pvalue = 2 * (n() / nrep))
```

The p-value of the difference between high and low valence is 0, which means we can reject the null hypothesis that there is not a difference between the mean popularities of songs that have high valence and low valence. The data provides convincing evidence that there is a difference in mean popularities of songs that have high valence versus low valence. 

### Popularity Difference Between Major and Minor
Null Hypothesis: There is not a difference between the mean popularities of songs that are in a major mode versus minor mode.
Alternative Hypothesis: There is a difference between the mean popularities of songs that are in a major mode versus minor mode.

```{r observed-difference-mode}
mean_mode <- songs %>%
  group_by(mode) %>%  
  summarise(samp_mean = mean(popularity)) %>%
  summarise(diff(samp_mean)) %>%
  pull()
mean_mode
```

```{r null-dist-mode}
null_dist_mode <- songs %>%
  specify(response = popularity, explanatory = mode) %>%
  hypothesize(null = "independence") %>%
  generate(reps = nrep, type = "permute") %>%
  calculate(stat = "diff in means", order = c("Major", "Minor"))
ggplot(data = null_dist_mode, aes(x = stat)) +
  geom_histogram(binwidth=.2) +
  geom_vline(xintercept = mean_mode, color = "red") +
  geom_vline(xintercept = -1 * mean_mode, color = "red") +
  labs(title = "Null distribution of differences in popularity means based on mode")
```

```{r p-value-mode}
null_dist_mode %>%
  filter(stat >= mean_mode) %>%
  summarise(pvalue = 2 * (n() / nrep))
```

The p-value of the difference between major and minor is 0, which means we can reject the null hypothesis that there is not a difference between the mean popularities of songs that are in a major or minor mode. The data provides convincing evidence that there is a difference in mean popularities of songs that are in a major mode versus a minor mode.  